---
output: github_document
---

[![Build Status](http://badges.herokuapp.com/travis/hypertidy/ncmeta?branch=master&env=BUILD_NAME=trusty_release&label=ubuntu)](https://travis-ci.org/hypertidy/ncmeta)
[![Build Status](http://badges.herokuapp.com/travis/hypertidy/ncmeta?branch=master&env=BUILD_NAME=osx_release&label=osx)](https://travis-ci.org/hypertidy/ncmeta) 
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/hypertidy/ncmeta?branch=master&svg=true)](https://ci.appveyor.com/project/mdsumner/ncmeta) 
[![CRAN status](http://www.r-pkg.org/badges/version/ncmeta)](https://cran.r-project.org/package=ncmeta)
[![](https://cranlogs.r-pkg.org/badges/ncmeta)](https://cran.r-project.org/package=ncmeta)
[![Coverage Status](https://img.shields.io/codecov/c/github/hypertidy/ncmeta/master.svg)](https://codecov.io/github/hypertidy/ncmeta?branch=master)

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# ncmeta

The `ncmeta` package provides straightforward NetCDF metadata, with a set of consistent entity-based functions for extracting metadata from a file or online source. We aim to fill a gap in between the generality and power of the NetCDF framework and ease of use. 

There are two main packages for using NetCDF in R, `RNetCDF` and `ncdf4` and ncmeta uses both, where appropriate. Both packages are very close the native API of NetCDF itself, and ncmeta simply provides an easier high-level interpretation. 

## About NetCDF

NetCDF is both a data model and an API, and provides a very general 
framework for expressing data formats. The explicit entities in NetCDF are 
**variables**, **dimensions** and **attributes** and ncmeta provides functions `nc_vars`, `nc_dims`, and `nc_atts` to extract their names, order and other metadata. There are matching functions `nc_var`, `nc_dim`, and `nc_att` with an extra identifier to extract specific information about an individual variable, dimension, or attribute. 

Also includes functions for implicit entities, these are **grids** and **axes**. These don't exist in the NetCDF specification explicitly, but are meaningful and worth making explicit. Many NetCDF tools don't explicitly present these concepts so grab hold of them with ncmeta!

A **grid** is an ordered set of dimensions, and the [Unidata site](https://www.unidata.ucar.edu/software/netcdf/netcdf/The-NetCDF-Data-Model.html#The-NetCDF-Data-Model) refers *informally* to this concept as **shape**. 

An **axis** is an *instance* of a dimension, the use of that dimension within a particular variable. 

These functions provide a more developer-friendly scheme for working with the range of formats provided by the NetCDF ecosystem. 

## Installation

Install ncmeta from CRAN with: 

```R
install.packages("ncmeta")
```

You can install the development version of ncmeta from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("hypertidy/ncmeta")
```

## Example

This example shows some of the functions for extracting information from a NetCDF source. 

```{r example,eval=FALSE}
library(ncmeta)
filename <- system.file("extdata", "S2008001.L3m_DAY_CHL_chlor_a_9km.nc", package = "ncmeta")
nc_inq(filename) # one-row summary of file

nc_dim(filename, 0)  ## first dimension
nc_dims(filename)  ## all dimensions
```


## Get involved!

Please let us know if you have any feedback, see the [Issues tab](https://github.com/hypertidy/ncmeta) if you found a bug or have a question. Feel free to email the maintainer directly for other questions. 


---

Please note that this project is released with a [Contributor Code of Conduct](https://github.com/hypertidy/ncmeta/blob/master/CODE_OF_CONDUCT.md#contributor-code-of-conduct).
  By participating in this project you agree to abide by its terms.
